A Plasma Torus Around a Young Low-Mass Star

### 2025-05-28 paper/ms.tex
update
ms width format

We observed TIC\,141146667 ($V$=16.2) for five hours on
17~February~2024 using the High Resolution Echelle Spectrometer
(HIRES; \citealt{vogt_hires_1994}) on the 10\,m Keck I telescope.  The
observations spanned the second half of the night, from 10:47 to 16:13
(UT). 
%The star's airmass spanned $z$=1.2-2.2, and we opted for a fixed 15
%minute cadence, except for a final 10 minute exposure due to
%increasing sky brightness at sunrise. 
\added{We opted for a fixed 15 minute cadence, and} used the C2 decker
(0$\farcs$86$\times$14$\farcs$0) in the red configuration, yielding a
resolution $R$$\approx$45{,}000 ($\delta
v$$\approx$6.7\,\kms).
%We binned the CCD readout by a factor of three in the spatial
%dimension, yielding $\approx$1,000 photons (S/N=33) per pixel in the
%continuum at 6500\,\AA. 
Strong winds contributed to 1$''$-1\farcs5 seeing, but conditions were
otherwise
favorable.  We reduced the echelleogram %to a one-dimensional spectrum
using the standard techniques of the California Planet Survey
\citep{Howard2010}.  
%Figure~\ref{fig:spec} shows the result in the
%vicinity of H$\alpha$ without any additional processing.

@@

In a follow-up study, \citet{Townsend2008} made synthetic light
curves, assuming that the optical depth scaled linearly with plasma
density.  They found that W-shaped eclipses, similar to those seen for
TIC\,141146667, can occur when the spin and magnetic axes are
moderately (15-45$^\circ$) misaligned.  The movies associated with
their work are available
online\footnote{\url{http://user.astro.wisc.edu/\~townsend/static.php?ref=rrm-movies\#Download\_Bundles}
last accessed 23 April 2025.}; for a $\omega/\omega_{\rm
c}$$\approx$0.5, edge-on system like TIC\,141146667, the photometric
eclipses have the correct shape, and the H$\alpha$ emission similarly
exhibits double-peaked behavior.  The plasma in this model accumulates
at antipodes 180$^\circ$ apart because the deepest local minima in the
effective potential exist along the line of intersection between the
rotational and magnetic planes (see Equation~22 of
\citealt{Townsend2005}).

@@

%The role of the star's magnetic field could be better understood through
%new observations, and new theory.  From the modeling perspective,
%there is a need for rigid-field magnetohydrodynamic models to
%go beyond previous work \citep{Townsend2008,Krticka2022}
%and to explore what field topologies explain the observed light
%curve shapes.  In particular, dynamo simulations of
%fully-convective M dwarfs have suggested that global-scale mean fields
%might be confined to a single hemisphere \citep{Brown2020}; such fields
%would yield accumulation surfaces quite different from those that have
%previously been explored.  Observationally, spectropolarimetry has the
%potential to assess both the field strength and topology.  An
%independent constraint on magnetic field topology may also come from
%radio observations, which have shown \citep{Kaur2024} that CPVs emit
%variable radio signals, including persistent and polarized components.
%Detecting radio emission produced by an electron cyclotron maser
%instability (e.g.~\citealt{Callingham2021}) in particular would provide
%a measurement of the field strength at the site of the emitting region,
%which could help clarify the magnetosphere's structure.

@@

{\it No Evidence For Binarity}---Any periodicity in the radial
velocity time-series is ruled out at the rotation period for
semi-amplitudes above 2.85\,\kms\ (at 3$\sigma$ confidence).  This
sets an upper limit on the mass of any putative companion on a four
hour orbit of $m \sin i $$<$2.4\,$M_{\rm Jup}$.  Regarding possible
companions at wider separations, the Gaia DR3 renormalized unit weight
error (RUWE) is 1.23, within the usual
range for apparently single sources.  There are no resolved companions
in the Gaia DR3 point source catalog.  Finally, we checked the TESS
light curve for evidence of secondary photometric periods by
subtracting the mean CPV signal over each sector and performing a
phase-dispersion minimization analysis
\citep{Stellingwerf1978,2021zndo...1011188B}.  There were no secondary
periods in the TESS data.  Previous work \citep{Bouma2024} has shown
that $\approx$30\% of CPVs show evidence for binarity from RUWE, and $\approx$40\% of CPVs show
evidence for binarity based on the presence of
secondary photometric periods.  This agrees with analyses showing that
multi-periodic low-mass stars are usually unresolved binaries
\citep{Tokovinin2018}.  Overall, the CPV binary fraction seems
consistent with that of field M dwarfs \citep{Winters2019}, pointing
to a weak or non-existent connection between the CPV phenomenon and
(wide) binarity.  For TIC\,141146667 specifically, we find no
evidence for stellar multiplicity. %, our data are minimally constraining
%for the scenario of a low-luminosity companion
%($F_1$/$F_2$$\lesssim$0.1) with apparent separation below 0$\farcs$1.

>>>>

We observed TIC\,141146667 ($V$=16.2) for five hours on
17~February~2024 using the High Resolution Echelle Spectrometer
(HIRES; \citealt{vogt_hires_1994}) on the 10\,m Keck I telescope.  The
observations spanned the second half of the night, from 10:47 to 16:13
(UT).  \added{We opted for a fixed 15 minute cadence, and} used the C2
decker (0$\farcs$86$\times$14$\farcs$0) in the red configuration,
yielding a resolution $R$$\approx$45{,}000 ($\delta
v$$\approx$6.7\,\kms).  Strong winds contributed to 1$''$-1\farcs5
seeing, but conditions were otherwise favorable.  We reduced the
echelleogram using the standard techniques of the California Planet
Survey \citep{Howard2010}.

@@

In a follow-up study, \citet{Townsend2008} made synthetic light
curves, assuming that the optical depth scaled linearly with plasma
density.  They found that W-shaped eclipses, similar to those seen for
TIC\,141146667, can occur when the spin and magnetic axes are
moderately (15-45$^\circ$) misaligned.  The movies associated with
their work are available
online\footnote{\url{http://user.astro.wisc.edu/\~townsend/static.php?ref=rrm-movies\#Download\_Bundles}
last accessed 23 April 2025.}; for a $\omega/\omega_{\rm
c}$$\approx$0.5 edge-on system like TIC\,141146667, the photometric
eclipses have the correct shape, and the H$\alpha$ emission similarly
exhibits double-peaked behavior.  The plasma clumps in this model are
180$^\circ$ apart because the deepest local minima in the effective
potential exist along the line of intersection between the rotational
and magnetic planes (see Equation~22 of \citealt{Townsend2005}).

@@

@@

{\it No Evidence For Binarity}---Any periodicity in the radial
velocity time-series is ruled out at the rotation period for
semi-amplitudes above 2.85\,\kms\ (at 3$\sigma$ confidence).  This
sets an upper limit on the mass of any putative companion on a four
hour orbit of $m \sin i $$<$2.4\,$M_{\rm Jup}$.  Regarding possible
companions at wider separations, the Gaia DR3 renormalized unit weight
error (RUWE) is 1.23, within the usual
range for apparently single sources.  There are no resolved companions
in the Gaia DR3 point source catalog.  Finally, we checked the TESS
light curve for evidence of secondary photometric periods by
subtracting the mean CPV signal over each sector and performing a
phase-dispersion minimization analysis
\citep{Stellingwerf1978,2021zndo...1011188B}.  There were no secondary
periods in the TESS data.  Previous work \citep{Bouma2024} has shown
that $\approx$30\% of CPVs show evidence for binarity from RUWE, and
$\approx$40\% of CPVs show evidence for binarity based on the presence
of secondary photometric periods.  This agrees with analyses showing
that multi-periodic low-mass stars are usually unresolved binaries
\citep{Tokovinin2018}.  Overall, the CPV binary fraction seems
consistent with that of field M dwarfs \citep{Winters2019}, pointing
to a weak or non-existent connection between the CPV phenomenon and
(wide) binarity.  For TIC\,141146667 specifically, we find no
evidence for stellar multiplicity.
